---
title: "CAMAS_TEST"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---
```{r include = FALSE}
# Load packages
library(tidyverse)
library(styler)
library(emmeans)
```
# Load the data

```{r include = FALSE}
# This loads the data
data_1 <- read.csv("/Users/christopheradlam/Desktop/Davis/R/GitHub Repos/Camas_Prairie/Data/CP_Camas_data.csv")
# table(data_1$NUMBER_CAMAS)
# data_1$NUMBER_CAMAS == 4
# data_1$PLOT_ID[data_1$NUMBER_CAMAS < 40]
```

# Subsetting the data (if necessary)
```{r include = FALSE}
# Subsetting the data
# data_1 <- data_1[-c(8, 19, 26, 27, 31, 32), ]
data_1 <- subset(data_1, TREATMENT == "C", select = c(PLOT_NUMBER, TREATMENT, NUMBER_CAMAS, PLOT_ID, AVG_LVS, AVG_FL, NUMBER_SCAPES))
data_1 <- subset(data_1, PLOT_NUMBER != 10, select = c(PLOT_NUMBER, TREATMENT, NUMBER_CAMAS, PLOT_ID, AVG_LVS, AVG_FL, NUMBER_SCAPES))

# better than subset maybe:
# data_1 %>% filter(NUMBER_CAMAS>200)
```

# Making plots to look for trends
```{r}
# Looking for relationships between above ground variables
plot(data_1$NUMBER_CAMAS, data_1$NUMBER_SCAPES)
abline(fit <- lm(data_1$NUMBER_SCAPES~data_1$NUMBER_CAMAS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_CAMAS, data_1$AVG_LVS)
abline(fit <- lm(data_1$AVG_LVS~data_1$NUMBER_CAMAS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_SCAPES, data_1$AVG_LVS)
abline(fit <- lm(data_1$AVG_LVS~data_1$NUMBER_SCAPES), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_CAMAS, data_1$AVG_FL)
abline(fit <- lm(data_1$AVG_FL~data_1$NUMBER_CAMAS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_SCAPES, data_1$AVG_FL)
abline(fit <- lm(data_1$AVG_FL~data_1$NUMBER_SCAPES), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$AVG_LVS, data_1$AVG_FL)
abline(fit <- lm(data_1$AVG_FL~data_1$AVG_LVS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

# Making a scatterplot matrix with p and r values in upper triangle
panel.cor <- function(x, y, digits = 2, cex.cor, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  # correlation coefficient
  r <- cor(x, y)
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste("r= ", txt, sep = "")
  text(0.5, 0.6, txt)

  # p-value calculation
  p <- cor.test(x, y)$p.value
  txt2 <- format(c(p, 0.123456789), digits = digits)[1]
  txt2 <- paste("p= ", txt2, sep = "")
  if (p < 0.01) txt2 <- paste("p= ", "<0.01", sep = "")
  text(0.5, 0.4, txt2)
}

lines <- function(x, y) {
  points(x, y, pch = ".")
  abline(lm(x~y), col = "red")
}

pairs(
  ~AVG_LVS + AVG_FL + NUMBER_SCAPES + NUMBER_CAMAS, data = data_1,
  main = "Simple Scatterplot Matrix", upper.panel = panel.cor, lower.panel = panel.smooth, panel = lines
)
```

```{r}
# control, unfiltered
data_1 <- read.csv("/Users/christopheradlam/Desktop/Davis/R/GitHub Repos/Camas_Prairie/Data/CP_Camas_data.csv")
data_1 <- subset(data_1, TREATMENT == "C", select = c(TREATMENT, NUMBER_CAMAS, PLOT_ID, AVG_LVS, AVG_FL, NUMBER_SCAPES, PLOT_NUMBER))

# filter?
data_1 <- data_1[-c(8, 19, 26, 27, 31, 32), ]


plot(data_1$NUMBER_CAMAS, data_1$NUMBER_SCAPES)
abline(fit <- lm(data_1$NUMBER_SCAPES~data_1$NUMBER_CAMAS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_CAMAS, data_1$AVG_LVS)
abline(fit <- lm(data_1$AVG_LVS~data_1$NUMBER_CAMAS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_SCAPES, data_1$AVG_LVS)
abline(fit <- lm(data_1$AVG_LVS~data_1$NUMBER_SCAPES), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_CAMAS, data_1$AVG_FL)
abline(fit <- lm(data_1$AVG_FL~data_1$NUMBER_CAMAS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$NUMBER_SCAPES, data_1$AVG_FL)
abline(fit <- lm(data_1$AVG_FL~data_1$NUMBER_SCAPES), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$AVG_LVS, data_1$AVG_FL)
abline(fit <- lm(data_1$AVG_FL~data_1$AVG_LVS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))
```

```{r}
# dig, changes over time after harvest
data_1 <- read.csv("/Users/christopheradlam/Desktop/Davis/R/GitHub Repos/Camas_Prairie/Data/CP_Camas_data.csv")
data_1 <- subset(data_1, TREATMENT == "T", select = c(TREATMENT, NUMBER_CAMAS, PLOT_ID, AVG_LVS, AVG_FL, NUMBER_SCAPES, YR_POST_HARVEST, PLOT_NUMBER, RATIO_FL))

# filter?
# data_1 <- data_1[-c(8, 19, 26, 27, 31, 32), ]
data_1 <- subset(data_1, PLOT_NUMBER != 10, select = c(PLOT_NUMBER, TREATMENT, NUMBER_CAMAS, PLOT_ID, AVG_LVS, AVG_FL, NUMBER_SCAPES, YR_POST_HARVEST, RATIO_FL))

plot(data_1$YR_POST_HARVEST, data_1$NUMBER_CAMAS)
abline(fit <- lm(data_1$NUMBER_CAMAS~data_1$YR_POST_HARVEST), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$YR_POST_HARVEST, data_1$AVG_LVS)
abline(fit <- lm(data_1$YR_POST_HARVEST~data_1$AVG_LVS), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$YR_POST_HARVEST, data_1$NUMBER_SCAPES)
abline(fit <- lm(data_1$YR_POST_HARVEST~data_1$NUMBER_SCAPES), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$YR_POST_HARVEST, data_1$AVG_FL)
abline(fit <- lm(data_1$YR_POST_HARVEST~data_1$AVG_FL), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))

plot(data_1$YR_POST_HARVEST, data_1$RATIO_FL)
abline(fit <- lm(data_1$YR_POST_HARVEST~data_1$RATIO_FL), col = "red")
legend("topright", bty = "n", legend = paste("R =", format(summary(fit)$r.squared, digits = 4)))
```

## Separating out the data from each variety into a vector. Want to analyze by year, compare T vs C, NUMBER_CAMAS, then NUMBER_SCAPES, then NUMBER_FLOWERS.
```{r}
# YR1, NUMBER_CAMAS.
YR1 <- subset(data_1, YR_POST_HARVEST == 1, select = c(TREATMENT, NUMBER_CAMAS, PLOT_ID))
CONTROL_YR1 <- YR1$NUMBER_CAMAS[YR1$TREATMENT == "C"]
DIG_YR1 <- YR1$NUMBER_CAMAS[YR1$TREATMENT == "T"]

# transform data?
# YR1$log_NUM <- log(YR1$NUMBER_CAMAS)

library(emmeans)
YR1$PLOT_ID <- as.character(YR1$PLOT_ID)
model_1 <- lm(NUMBER_CAMAS ~ TREATMENT, data = YR1) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR1, aes(x = TREATMENT, y = NUMBER_CAMAS)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR1, DIG_YR1, paired = T)
```


```{r}
# YR1, NUMBER_SCAPES
YR1 <- subset(data_1, YR_POST_HARVEST == 1, select = c(TREATMENT, NUMBER_SCAPES, PLOT_ID))
CONTROL_YR1 <- YR1$NUMBER_SCAPES[YR1$TREATMENT == "C"]
DIG_YR1 <- YR1$NUMBER_SCAPES[YR1$TREATMENT == "T"]

library(emmeans)
# YR1$NUMBER_SCAPES <- log(YR1$NUMBER_SCAPES)
YR1$PLOT_ID <- as.character(YR1$PLOT_ID)
model_1 <- lm(NUMBER_SCAPES ~ TREATMENT, data = YR1) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR1, aes(x = TREATMENT, y = NUMBER_SCAPES)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR1, DIG_YR1, paired = T)
```
```{r}
# YR1, RATIO_FL
YR1 <- subset(data_1, YR_POST_HARVEST == 1, select = c(TREATMENT, RATIO_FL, PLOT_ID))
CONTROL_YR1 <- YR1$RATIO_FL[YR1$TREATMENT == "C"]
DIG_YR1 <- YR1$RATIO_FL[YR1$TREATMENT == "T"]

library(emmeans)
# YR1$NUMBER_SCAPES <- log(YR1$NUMBER_SCAPES)
YR1$PLOT_ID <- as.character(YR1$PLOT_ID)
model_1 <- lm(RATIO_FL ~ TREATMENT, data = YR1) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR1, aes(x = TREATMENT, y = RATIO_FL)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR1, DIG_YR1, paired = T)
```

```{r}
# YR2, RATIO_FL
YR2 <- subset(data_1, YR_POST_HARVEST == 2, select = c(TREATMENT, RATIO_FL, PLOT_ID))
CONTROL_YR2 <- YR1$RATIO_FL[YR2$TREATMENT == "C"]
DIG_YR2 <- YR1$RATIO_FL[YR2$TREATMENT == "T"]

library(emmeans)
# YR1$NUMBER_SCAPES <- log(YR1$NUMBER_SCAPES)
YR2$PLOT_ID <- as.character(YR2$PLOT_ID)
model_1 <- lm(RATIO_FL ~ TREATMENT, data = YR2) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR2, aes(x = TREATMENT, y = RATIO_FL)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR2, DIG_YR2, paired = T)
```

```{r}
# YR2, RATIO_FL
YR3 <- subset(data_1, YR_POST_HARVEST == 3, select = c(TREATMENT, RATIO_FL, PLOT_ID))
CONTROL_YR3 <- YR3$RATIO_FL[YR3$TREATMENT == "C"]
DIG_YR3 <- YR3$RATIO_FL[YR3$TREATMENT == "T"]

library(emmeans)
# YR1$NUMBER_SCAPES <- log(YR1$NUMBER_SCAPES)
YR3$PLOT_ID <- as.character(YR3$PLOT_ID)
model_1 <- lm(RATIO_FL ~ TREATMENT, data = YR3) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR3, aes(x = TREATMENT, y = RATIO_FL)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR3, DIG_YR3, paired = T)
```


```{r}
# YR2, NUMBER_CAMAS
YR2 <- subset(data_1, YR_POST_HARVEST == 2, select = c(TREATMENT, NUMBER_CAMAS, PLOT_ID))
CONTROL_YR2 <- YR2$NUMBER_CAMAS[YR2$TREATMENT == "C"]
DIG_YR2 <- YR2$NUMBER_CAMAS[YR2$TREATMENT == "T"]

library(emmeans)
YR2$PLOT_ID <- as.character(YR2$PLOT_ID)
model_1 <- lm(NUMBER_CAMAS ~ TREATMENT, data = YR2) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR2, aes(x = TREATMENT, y = NUMBER_CAMAS)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR2, DIG_YR2, paired = T)
```

```{r}
# YR2, NUMBER_SCAPES
YR2 <- subset(data_1, YR_POST_HARVEST == 2, select = c(TREATMENT, NUMBER_SCAPES, PLOT_ID))
CONTROL_YR2 <- YR2$NUMBER_SCAPES[YR2$TREATMENT == "C"]
DIG_YR2 <- YR2$NUMBER_SCAPES[YR2$TREATMENT == "T"]

library(emmeans)
YR2$PLOT_ID <- as.character(YR2$PLOT_ID)
model_1 <- lm(NUMBER_SCAPES ~ TREATMENT, data = YR2) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR2, aes(x = TREATMENT, y = NUMBER_SCAPES)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR2, DIG_YR2, paired = T)
```

```{r}
# YR3, NUMBER_CAMAS
YR3 <- subset(data_1, YR_POST_HARVEST == 3, select = c(TREATMENT, NUMBER_CAMAS, PLOT_ID))
CONTROL_YR3 <- YR3$NUMBER_CAMAS[YR3$TREATMENT == "C"]
DIG_YR3 <- YR3$NUMBER_CAMAS[YR3$TREATMENT == "T"]

library(emmeans)
YR3$PLOT_ID <- as.character(YR3$PLOT_ID)
model_1 <- lm(NUMBER_CAMAS ~ TREATMENT, data = YR3) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR3, aes(x = TREATMENT, y = NUMBER_CAMAS)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR3, DIG_YR3, paired = T)
```

```{r}
# YR3, NUMBER_SCAPES
YR3 <- subset(data_1, YR_POST_HARVEST == 3, select = c(TREATMENT, NUMBER_SCAPES, PLOT_ID))
CONTROL_YR3 <- YR3$NUMBER_SCAPES[YR3$TREATMENT == "C"]
DIG_YR3 <- YR3$NUMBER_SCAPES[YR3$TREATMENT == "T"]

library(emmeans)
# YR3$NUMBER_SCAPES <- log(YR3$NUMBER_SCAPES)
YR3$PLOT_ID <- as.character(YR3$PLOT_ID)
model_1 <- lm(NUMBER_SCAPES ~ TREATMENT, data = YR3) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(YR3, aes(x = TREATMENT, y = NUMBER_SCAPES)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR3, DIG_YR3, paired = T)
```

```{r}
# Mean number of leaves
# YR1
LVS <- subset(data_1, YR_POST_HARVEST == 1, select = c(TREATMENT, AVG_LVS, PLOT_ID))
CONTROL_YR1 <- LVS$AVG_LVS[LVS$TREATMENT == "C"]
DIG_YR1 <- LVS$AVG_LVS[LVS$TREATMENT == "T"]

library(emmeans)
# YR3$NUMBER_SCAPES <- log(YR3$NUMBER_SCAPES)
LVS$PLOT_ID <- as.character(LVS$PLOT_ID)
model_1 <- lm(AVG_LVS ~ TREATMENT, data = LVS) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(LVS, aes(x = TREATMENT, y = AVG_LVS)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR1, DIG_YR1, paired = T)
```
```{r}
# Mean number of leaves
# YR2
LVS <- subset(data_1, YR_POST_HARVEST == 2, select = c(TREATMENT, AVG_LVS, PLOT_ID))
CONTROL_YR2 <- LVS$AVG_LVS[LVS$TREATMENT == "C"]
DIG_YR2 <- LVS$AVG_LVS[LVS$TREATMENT == "T"]

library(emmeans)
# YR3$NUMBER_SCAPES <- log(YR3$NUMBER_SCAPES)
LVS$PLOT_ID <- as.character(LVS$PLOT_ID)
model_1 <- lm(AVG_LVS ~ TREATMENT, data = LVS) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(LVS, aes(x = TREATMENT, y = AVG_LVS)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR2, DIG_YR2, paired = T)
```

```{r}
# Mean number of leaves, YR3
LVS <- subset(data_1, YR_POST_HARVEST == 3, select = c(TREATMENT, AVG_LVS, PLOT_ID))
CONTROL_YR3 <- LVS$AVG_LVS[LVS$TREATMENT == "C"]
DIG_YR3 <- LVS$AVG_LVS[LVS$TREATMENT == "T"]

library(emmeans)
# YR3$NUMBER_SCAPES <- log(YR3$NUMBER_SCAPES)
LVS$PLOT_ID <- as.character(LVS$PLOT_ID)
model_1 <- lm(AVG_LVS ~ TREATMENT, data = LVS) # fit the model

# model with treatment differences
differences_model1 <- emmeans(model_1, pairwise~TREATMENT) # `pairwise~Variety` means calculate all pairwise differences among varieties
summary(differences_model1, level = 0.90, infer = c(T, F))$contrast

# Model diagnostic plots
library(ggplot2)
ggplot(LVS, aes(x = TREATMENT, y = AVG_LVS)) + geom_boxplot() + geom_jitter(width = .2)

par(mfrow = c(1, 2)) # necessary to get two plots next to each other
plot(model_1, which = c(2, 5)) # there are several diagnostic plots that R can make. For simple experiments, these two are good.

# t-test
t.test(CONTROL_YR3, DIG_YR3, paired = T)
```


## Calculate the power of the test
```{r}
power.t.test(n = 12, power = 0.9, sd = 4, sig.level = 0.1)
```
```{r}
# Plots
data_1 <- subset(data_1, PLOT_NUMBER != 10, select = c(PLOT_NUMBER, TREATMENT, NUMBER_CAMAS, PLOT_ID, AVG_LVS, AVG_FL, NUMBER_SCAPES, YR_POST_HARVEST))

## Number Camas in each treatment
TRT <- subset(data_1, TREATMENT == "T", select = c(NUMBER_CAMAS, YR_POST_HARVEST))
plot(TRT$YR_POST_HARVEST, TRT$NUMBER_CAMAS)

# number scapes in each treatment
TRT <- subset(data_1, TREATMENT == "C", select = c(NUMBER_SCAPES, YR_POST_HARVEST))
plot(TRT$YR_POST_HARVEST, TRT$NUMBER_SCAPES)

# AVG NUMBER FLOWERS
TRT <- subset(data_1, TREATMENT == "T", select = c(AVG_LVS, YR_POST_HARVEST))
plot(TRT$YR_POST_HARVEST, TRT$AVG_LVS)
abline(lm(TRT$AVG_LVS~TRT$YR_POST_HARVEST))

TRT <- subset(data_1, TREATMENT == "C", select = c(AVG_LVS, YR_POST_HARVEST))
plot(TRT$YR_POST_HARVEST, TRT$AVG_LVS)
abline(lm(TRT$AVG_LVS~TRT$YR_POST_HARVEST))

DIG_YR1 <- TRT$AVG_LVS[TRT$YR_POST_HARVEST == "1"]
DIG_YR2 <- TRT$AVG_LVS[TRT$YR_POST_HARVEST == "2"]
DIG_YR3 <- TRT$AVG_LVS[TRT$YR_POST_HARVEST == "3"]

CTL <- subset(data_1, TREATMENT == "C", select = c(AVG_LVS, YR_POST_HARVEST))
CONTROL <- CTL$AVG_LVS
CONTROL_YR1 <- CTL$AVG_LVS[CTL$YR_POST_HARVEST == "1"]
CONTROL_YR2 <- CTL$AVG_LVS[CTL$YR_POST_HARVEST == "2"]
CONTROL_YR3 <- CTL$AVG_LVS[CTL$YR_POST_HARVEST == "3"]

boxplot(CONTROL_YR1, CONTROL_YR2, CONTROL_YR3)
boxplot(DIG_YR1, DIG_YR2, DIG_YR3, CONTROL_YR1, CONTROL_YR2, CONTROL_YR3)
```



# Plots----------------
## Main data
```{r}
#NUMBER_CAMAS
# Diagnostic plots
## NOT NORMAL
op = par(mfrow=c(1,2))
plot(camas.mod.NUMBER_CAMAS,which = c(2,3))

# blox plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = NUMBER_CAMAS)) + geom_boxplot() + geom_jitter(width = 0.2)

# violin plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = NUMBER_CAMAS)) + geom_violin(bw = 5) + geom_jitter(width = 0.2)
```

```{r}
#NUMBER_SCAPES
# Diagnostic plots
## Normality looks good; some heterogeneity of variances
op = par(mfrow=c(1,2))
plot(camas.mod.NUMBER_SCAPES,which = c(2,3))

# blox plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = NUMBER_SCAPES)) + geom_boxplot() + geom_jitter(width = 0.2)

# violin plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = NUMBER_SCAPES)) + geom_violin(bw = 5) + geom_jitter(width = 0.2)
```

```{r}
# AVG_LVS
# Diagnostic plots
## looks good
op = par(mfrow=c(1,2))
plot(camas.mod.AVG_LVS,which = c(2,3))

# blox plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = AVG_LVS)) + geom_boxplot() + geom_jitter(width = 0.2)

# violin plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = AVG_LVS)) + geom_violin(bw = 0.5) + geom_jitter(width = 0.2)
```

```{r}
# AVG_FL
# Diagnostic plots
## Looks good
op = par(mfrow=c(1,2))
plot(camas.mod.AVG_FL,which = c(2,3))

# blox plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = AVG_FL)) + geom_boxplot() + geom_jitter(width = 0.2)

# violin plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = AVG_FL)) + geom_violin(bw = 1.5) + geom_jitter(width = 0.2)
```

```{r}
# RATIO_FL
# Diagnostic plots
# NOT normally distributed because of NUMBER_CAMAS
op = par(mfrow=c(1,2))
plot(camas.mod.NUMBER_CAMAS,which = c(2,3))
# blox plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = RATIO_FL)) + geom_boxplot() + geom_jitter(width = 0.2)

# violin plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = RATIO_FL)) + geom_violin(bw = 0.5) + geom_jitter(width = 0.2)
```
## Difference between paired C and T, by year

#ANOVA-----------------
## Main data
```{r}
camas.data <- data_1 %>%
  select(NUMBER_CAMAS, NUMBER_SCAPES, AVG_LVS, AVG_FL, TREATMENT, PLOT_ID, YR_POST_HARVEST, RATIO_FL, PLOT_NUMBER) %>%
  mutate(YR_TRT = paste(TREATMENT, YR_POST_HARVEST))

# Not significant: NUMBER_CAMAS and AVG_FL
camas.mod.NUMBER_CAMAS <- lm(NUMBER_CAMAS ~ YR_TRT, data = camas.data)
anova(camas.mod.NUMBER_CAMAS)

camas.mod.AVG_FL <- lm(AVG_FL ~ YR_TRT, data = camas.data)
anova(camas.mod.AVG_FL)

# Significant: NUMBER_SCAPES, AVG_LVS, RATIO_FL
camas.mod.NUMBER_SCAPES <- lm(NUMBER_SCAPES ~ YR_TRT, data = camas.data)
anova(camas.mod.NUMBER_SCAPES)

camas.mod.AVG_LVS <- lm(AVG_LVS ~ YR_TRT, data = camas.data)
anova(camas.mod.AVG_LVS)

camas.mod.RATIO_FL <- lm(RATIO_FL ~ YR_TRT, data = camas.data)
anova(camas.mod.RATIO_FL)
```

## Difference between paired C and T, by year
### Set up the paired differences
```{r}
# IF PLOT_ID in one row = PLOT_ID in another row, THEN subtract the response value from the T row from the C row. How???

## The following doesn't work because plots don't necessarily add up.
## YR1
camas <- data_1 %>% 
  filter(YR_POST_HARVEST == 1) %>% 
  select(PLOT_ID, TREATMENT, PLOT_NUMBER, NUMBER_CAMAS, AVG_FL, NUMBER_SCAPES, RATIO_FL, AVG_LVS) 

PLOT_NUM_DIFF <- camas$PLOT_NUMBER[1:11]
NUM_CAMAS_C <- camas$NUMBER_CAMAS[1:11] 
NUM_CAMAS_T <- camas$NUMBER_CAMAS[12:22]
NUM_SCAPES_C <- camas$NUMBER_SCAPES[1:11]
NUM_SCAPES_T <- camas$NUMBER_SCAPES[12:22]
AVG_LVS_C <- camas$AVG_LVS[1:11]
AVG_LVS_T <- camas$AVG_LVS[12:22]
AVG_FL_C <- camas$AVG_FL[1:11]
AVG_FL_T <- camas$AVG_FL[12:22]  
RATIO_FL_C <- camas$AVG_FL[1:11]
RATIO_FL_T <- camas$AVG_FL[12:22] 

NUM_CAMAS_DIFF1 <- NUM_CAMAS_C - NUM_CAMAS_T
NUM_SCAPES_DIFF1 <- NUM_SCAPES_C - NUM_SCAPES_T
AVG_LVS_DIFF1 <- AVG_LVS_C - AVG_LVS_T
AVG_FL_DIFF1 <- AVG_FL_C - AVG_FL_T
RATIO_FL_DIFF1 <- RATIO_FL_C - RATIO_FL_T

## YR2
camas <- data_1 %>% 
  filter(YR_POST_HARVEST == 2) %>% 
  select(PLOT_ID, TREATMENT, PLOT_NUMBER, NUMBER_CAMAS, AVG_FL, NUMBER_SCAPES, RATIO_FL, AVG_LVS) %>% 
  arrange(TREATMENT)

PLOT_NUM_DIFF <- camas$PLOT_NUMBER[1:11]
NUM_CAMAS_C <- camas$NUMBER_CAMAS[1:11] 
NUM_CAMAS_T <- camas$NUMBER_CAMAS[12:22]
NUM_SCAPES_C <- camas$NUMBER_SCAPES[1:11]
NUM_SCAPES_T <- camas$NUMBER_SCAPES[12:22]
AVG_LVS_C <- camas$AVG_LVS[1:11]
AVG_LVS_T <- camas$AVG_LVS[12:22]
AVG_FL_C <- camas$AVG_FL[1:11]
AVG_FL_T <- camas$AVG_FL[12:22]  
RATIO_FL_C <- camas$AVG_FL[1:11]
RATIO_FL_T <- camas$AVG_FL[12:22] 

NUM_CAMAS_DIFF2 <- NUM_CAMAS_C - NUM_CAMAS_T
NUM_SCAPES_DIFF2 <- NUM_SCAPES_C - NUM_SCAPES_T
AVG_LVS_DIFF2 <- AVG_LVS_C - AVG_LVS_T
AVG_FL_DIFF2 <- AVG_FL_C - AVG_FL_T
RATIO_FL_DIFF2 <- RATIO_FL_C - RATIO_FL_T

## YR3
camas <- data_1 %>% 
  filter(YR_POST_HARVEST == 3) %>% 
  select(PLOT_ID, TREATMENT, PLOT_NUMBER, NUMBER_CAMAS, AVG_FL, NUMBER_SCAPES, RATIO_FL, AVG_LVS) 

PLOT_NUM_DIFF <- camas$PLOT_NUMBER[1:9]
NUM_CAMAS_C <- camas$NUMBER_CAMAS[1:9] 
NUM_CAMAS_T <- camas$NUMBER_CAMAS[10:18]
NUM_SCAPES_C <- camas$NUMBER_SCAPES[1:9]
NUM_SCAPES_T <- camas$NUMBER_SCAPES[10:18]
AVG_LVS_C <- camas$AVG_LVS[1:9]
AVG_LVS_T <- camas$AVG_LVS[10:18]
AVG_FL_C <- camas$AVG_FL[1:9]
AVG_FL_T <- camas$AVG_FL[10:18]  
RATIO_FL_C <- camas$AVG_FL[1:9]
RATIO_FL_T <- camas$AVG_FL[10:18] 

NUM_CAMAS_DIFF3 <- NUM_CAMAS_C - NUM_CAMAS_T
NUM_SCAPES_DIFF3 <- NUM_SCAPES_C - NUM_SCAPES_T
AVG_LVS_DIFF3 <- AVG_LVS_C - AVG_LVS_T
AVG_FL_DIFF3 <- AVG_FL_C - AVG_FL_T
RATIO_FL_DIFF3 <- RATIO_FL_C - RATIO_FL_T

diff <- data.frame(PLOT_NUM_DIFF, NUM_CAMAS_DIFF1, NUM_SCAPES_DIFF1, AVG_LVS_DIFF1, AVG_FL_DIFF1, RATIO_FL_DIFF1, NUM_CAMAS_DIFF2, NUM_SCAPES_DIFF2, AVG_LVS_DIFF2, AVG_FL_DIFF2, RATIO_FL_DIFF2, NUM_CAMAS_DIFF3, NUM_SCAPES_DIFF3, AVG_LVS_DIFF3, AVG_FL_DIFF3, RATIO_FL_DIFF3)

```
```{r}
diff_mod <- lm(NUM_CAMAS_DIFF ~ YEAR, diff)
```

### testing differences between paired plots, YR1

```{r}
# plot
plot(diff_YR1)
ggplot(NUM_CAMAS_DIFF, mapping = aes(x = YR_TRT, y = AVG_FL)) + geom_boxplot() + geom_jitter(width = 0.2)

# violin plot
ggplot(camas.data, mapping = aes(x = YR_TRT, y = AVG_FL)) + geom_violin(bw = 1.5) + geom_jitter(width = 0.2)




# model: number camas
diff_mod_CAMAS <- lm(NUM_CAMAS_DIFF ~ TREATMENT, diff_YR1)
op = par(mfrow=c(1,2))
plot(diff_mod_CAMAS,which = c(2,3))

diff_mod_CAMAS

# model: number scapes
diff_mod_SCAPES <- lm(NUM_SCAPES_DIFF ~ PLOT_NUM_DIFF, diff_YR1)
op = par(mfrow=c(1,2))
plot(diff_mod_SCAPES,which = c(2,3))

# model: number camas
diff_mod_LVS <- lm(AVG_LVS_DIFF ~ PLOT_NUM_DIFF, diff_YR1)
op = par(mfrow=c(1,2))
plot(diff_mod_LVS,which = c(2,3))

# model: number camas
diff_mod_AVG_FL <- lm(AVG_FL_DIFF ~ PLOT_NUM_DIFF, diff_YR1)
op = par(mfrow=c(1,2))
plot(diff_mod_AVG_FL ,which = c(2,3))

# model: number camas
diff_mod_RATIO_FL <- lm(RATIO_FL_DIFF ~ PLOT_NUM_DIFF, diff_YR1)
op = par(mfrow=c(1,2))
plot(diff_mod_RATIO_FL,which = c(2,3))


```

```{r}
diff_camas = emmeans(diff_mod_CAMAS,pairwise~NUM_CAMAS_DIFF)
summary(diff_camas,level = 0.95,infer = c(T,T))$contrast

diff_camas = emmeans(diff_mod_SCAPES,pairwise~Drug)
summary(diff_camas,level = 0.95,infer = c(T,T))$contrast

diff_camas = emmeans(diff_mod_LVS,pairwise~Drug)
summary(diff_camas,level = 0.95,infer = c(T,T))$contrast

diff_camas = emmeans(diff_mod_AVG_FL,pairwise~Drug)
summary(diff_camas,level = 0.95,infer = c(T,T))$contrast

diff_camas = emmeans(diff_mod_RATIO_FL,pairwise~Drug)
summary(diff_camas,level = 0.95,infer = c(T,T))$contrast

```


#t-tests
```{r}
# NUMBER_SCAPES : significant:
all_differences <- emmeans(camas.mod.NUMBER_SCAPES, pairwise~YR_TRT)
contrast(all_differences)
as.data.frame(contrast(all_differences)$contrasts) %>%
  filter(p.value < 0.05)

# AVG_LVS : significant:
all_differences <- emmeans(camas.mod.AVG_LVS, pairwise~YR_TRT)
contrast(all_differences)
as.data.frame(contrast(all_differences)$contrasts) %>%
  filter(p.value < 0.05)
```

```{r}

```

